---
title: "Evaluating the Role of Airbnb in Zurich’s Neighbourhoods"
author: "Ana Mas Urquijo, Azeglio Martinelli"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    number_sections: true
    toc: true               # table of contents
    toc_depth: 3            # heading levels displayed
    toc_float:
      collapsed: false      # toc in panel on the left
      smooth_scroll: true   # scroll instead of jump
---

```{r knitr-setup, include = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,       # do not display R code
  include = TRUE,     # display R result
  message = FALSE,
  warning = FALSE,
  error = FALSE
  
  )
```

# Introduction

Zurich is widely recognised as one of the most constrained housing markets in Switzerland. Strong population growth, limited land availability, and a persistently low vacancy rate have made access to long-term rental housing increasingly difficult. For many residents, the process of finding an apartment involves long waiting lists and crowded viewings, reflecting the intense competition for available housing. Ensuring sufficient long-term housing supply has therefore become a key policy challenge for the city.

At the same time, short-term rental platforms such as Airbnb have expanded significantly in Zurich over the past decade. These platforms offer flexible accommodation options for tourists and business travellers and can provide additional income opportunities for residents. However, when residential units are offered frequently or on a year-round basis as short-term rentals, they may reduce the availability of housing for long-term residents. This potential trade-off between tourism demand and housing supply has sparked increasing public debate.

In response to similar challenges, several Swiss cities and cantons have introduced regulatory measures aimed at limiting the intensity of short-term rentals. One commonly discussed instrument is a cap on the number of nights per year that a dwelling may be rented to tourists, such as a limit of 90 nights annually. The City of Zurich is currently considering whether comparable regulations would be appropriate and how such measures could be implemented in a targeted and proportionate way.

Importantly, the role of Airbnb is unlikely to be uniform across the city. In some neighbourhoods, short-term rentals may complement existing accommodation infrastructure, particularly where hotel capacity is limited. In other areas, a high concentration of listings or hosts managing multiple properties may indicate a more commercial use of residential housing. At the same time, guest reviews and booking activity provide insight into whether Airbnb listings deliver positive experiences for visitors, which is a relevant consideration for tourism policy.

This analysis is conducted for a hypothetical client, the City of Zurich, with the objective of evaluating the role of Airbnb across Zurich neighbourhoods. Using publicly available data, the study examines spatial patterns of listings, host concentration, rental intensity, and guest experiences. The overarching goal is to assess where short-term rentals appear to complement the city’s housing and tourism landscape and where targeted regulatory measures may be justified.

# Data Sources

The analysis is based on publicly available data from **Inside Airbnb** ([https://insideairbnb.com/get-the-data/](https://insideairbnb.com/get-the-data/)), an independent, non-commercial project that provides detailed information on Airbnb listings in cities around the world. The platform compiles data directly from Airbnb’s website and aims to support research and policy discussions around short-term rentals.

Two datasets for the city of Zurich are used in this study:

1. **Listings dataset**
   This dataset contains detailed information on individual Airbnb listings, including their location, room type, price, availability, minimum stay requirements, and host characteristics. It also includes variables such as the number of listings per host and whether a listing is associated with a license.

2. **Reviews dataset**
   This dataset contains information on guest reviews, including review dates and the number of reviews over time. It allows for an assessment of demand dynamics and guest experiences across listings and neighbourhoods.

The two datasets are merged using the unique listing identifier. This combined dataset enables a joint analysis of spatial distribution, host behaviour, usage intensity, and guest feedback. All data preparation steps, including cleaning and merging, are documented and highlighted in the analysis to ensure full reproducibility.


# Objectives and Research Questions

The main objective of this analysis is to support the City of Zurich in evaluating whether, and where, regulatory measures on short-term rentals may be appropriate. Rather than treating Airbnb as a uniform phenomenon, the analysis focuses on neighbourhood-level differences and host behaviour.

Specifically, the study addresses the following research questions:

**RQ1:**
How are Airbnb listings distributed across Zurich neighbourhoods, and which areas show the highest concentration of short-term rentals?

**RQ2:**
Are Airbnb listings more prevalent in certain neighbourhoods, suggesting that short-term rentals may complement local accommodation needs rather than displace long-term housing?

**RQ3:**
To what extent is Airbnb activity in Zurich driven by a small number of hosts offering multiple listings, indicating a more commercial use of residential properties?

**RQ4:**
Which neighbourhoods exhibit high short-term rental intensity, as reflected by listing availability throughout the year and minimum stay requirements?

**RQ5:**
Do guest reviews suggest that Airbnb listings in Zurich provide positive tourist experiences, and does this vary across neighbourhoods or host types?

**RQ6:**
Based on listing concentration, host behaviour, and usage intensity, which neighbourhoods would be most affected by potential regulatory measures on short-term rentals?

Together, these questions provide a structured framework for assessing the role of Airbnb in Zurich and for identifying areas where targeted regulation may be most relevant.


# Data Preparation

The cell below was run only once to unpack the reviews data from a .gz
compressed folder. The code cell unpacked the data and directly deleted
the compressed files.

```{r, eval=FALSE}
library(R.utils)
gunzip("Data/reviews.csv.gz")
```

First, the listings and reviews datasets are imported from CSV files using read_csv and stored as d.listings and d.reviews. After loading the listings data, the amount of missing data is examined by calculating both the absolute number of missing values in each column and the proportion of missing values relative to the total number of observations. This provides an overview of data completeness and helps identify variables that may require cleaning or removal.

Based on this inspection, the column license is removed from the listings dataset because it contains only missing values and therefore provides no usable information for the analysis. The remaining missing values in selected numeric variables are then addressed by replacing NA values with zeros for price and reviews_per_month, ensuring that these variables contain valid numeric values and can be used in subsequent analyses. The last_review variable is intentionally left unchanged, as imputing a date where none exists would not be meaningful.

The same missing-data diagnostics are then applied to the reviews dataset by again calculating the number and proportion of missing values in each column. Missing values in the comments column are replaced with the string "Unknown" so that all review records contain a valid text entry and can be processed consistently. Afterwards, all non-English comments are deleted. It is explained in the next section why this was done.

Finally, the listings dataset is checked for duplicate listing identifiers. This is done by counting how often each id appears, filtering for those that occur more than once, and displaying only the distinct duplicated entries. This step verifies whether the primary key in the listings data is unique, which is important for reliable joins and further analysis.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(cld3)

d.listings <- read_csv("Data/listings.csv")
d.reviews <- read_csv("Data/reviews.csv")

# missing data in each column
colSums(is.na(d.listings))
# percentage of missing data
colMeans(is.na(d.listings))

# drop column license as it contains only NA values
d.listings <- d.listings %>% 
  select(-license)

# replace NA's in other columns so that we have valid values for analysis (except for last_review because here it does not make sense to enter a somewhat valid date for NA values)
d.listings <- d.listings %>% 
  replace_na(list(price = 0, reviews_per_month = 0))

# same checks for reviews and replace NA's
colSums(is.na(d.reviews))
colMeans(is.na(d.reviews))

d.reviews <- d.reviews %>% 
  replace_na(list(comments = 'Unkown'))

# filter out non-English comments
d.reviews_eng <- d.reviews %>%
  mutate(
    detected_language = detect_language(comments)
  ) %>%
  filter(
    detected_language == "en" | is.na(detected_language)
  ) %>%
  select(-detected_language)

# Check for duplicate Id's in listings
d.listings %>%
  add_count(id) %>%
  filter(n>1) %>%
  distinct()

```

In the next cell below, the two data sets were merged.

First, a join condition is explicitly defined to link the two datasets by matching the id column in the listings data with the listing_id column in the reviews data. Using this condition, the cleaned listings dataset is merged with the reviews dataset using a left join, ensuring that all listings are retained in the merged dataset while review information is added where a matching review exists.

After performing the merge, sanity checks are carried out to assess whether the join behaved as expected. The total number of rows in the merged dataset is computed and compared with the number of rows in the listings data. In addition, the number of unmatched rows is calculated by counting how many entries have missing values in the review identifier column, indicating listings for which no corresponding review record was found.

To further investigate unmatched cases, an anti-join is performed between the listings and reviews datasets. This operation returns only those listings that do not have a matching review. The result is inspected manually using the View() function, allowing a detailed examination of all columns. This inspection confirms that the unmatched entries correspond to listings without any reviews, which explains the absence of a matching listing_id in the reviews data, particularly given that the last_review field contains only missing values for these cases.

Finally, to improve clarity and avoid ambiguity in the merged dataset, review-related columns are renamed. The review identifier column originating from the reviews table is renamed to review_id, and the review date column is renamed to review_date, making the structure and meaning of the merged data clearer for subsequent analysis.


```{r}
by <- join_by(id == listing_id)

d.merged <- d.listings %>%
  left_join(d.reviews_eng, by = by)

# Sanity checks for merge below

# Check number of matches where join columns of reviews are NA
d.merged %>%
  summarise(
    total = n(),
    unmatched = sum(is.na(id.y))
  )

# anti_join to inspect any mismatches
anti_join(d.listings, d.reviews, by = c("id" = "listing_id")) |>
  View()

# Upon closer, manual inspection wiht the View() it is evident that all the non matched columns exist because they do not have a review and therefore no listing_id to match (because column last_review is all NA's).

# rename different review columns
d.merged <- d.merged %>% 
  rename(review_id = id.y, review_date = date)

```


# Chapter of choice / Sentiment Analysis

For the chapter of choice, a Sentiment Analysis on the comment is done to augment the merged dataset with a sentiment score. This is the reason why all non-English comments were deleted in the Data Preparation section because the Sentiment Analysis employed can only handle English text and all non-English text is assigned with negative scores even if the comment itself is very positive.

First, we start be creating a reduced data frame that contains only the review identifier and the associated review text. Then, the following text preprocessing steps are done on this comments data frame. All comments are converted to lowercase to ensure that words differing only by capitalization are treated as the same token during analysis. After that, all punctuation is removed so that it does no interfere with word matching in the sentiment lexicon.

With the preprocessing done, the comments are tokenized. Meaning each comment is first split into individiual words, with one word per row, to allow for easier analysis of the data frame. Of course, review_id is kept to later aggregate back to comment level. 

The Bing sentiment lexicon is then loaded. With this lexicon, the tokenized words are catgorized in a binary fashion, either positive or negative. The tokenized words are joined with this lexicon so that only words appearing in the sentiment dictionary are retained and assigned a corresponding sentiment. Next, the sentiment information is aggregated at the comment level. The categorization of each word is safed into one of two columns, either 'n_positive' or 'n_negative', as counts. These counts are then assigned a score of 1 if the comment contains more positive than negative words, -1 for the opposite case and 0 if the numbers are equal, indicating a neutral comment. 

Lastly, the sentiment score columns are merged with the original 'd.merged' dataset on the review identifier using a left join. Any reviews that did not contain words found in the bing lexicon are assigned a neutral score of 0. This is done so that the score column contains no empty values. The result is the original dataset being augmented with a sentiment score for each review that reflects its overall tone and can be used for visualizations.

```{r}

# this is here just so that the cell can be re-run without duplicating the new columns
d.merged <- d.merged %>% 
  select(c(id:comments))

library(tidytext)
library(textdata)

comments_df <- d.merged %>%
  select(review_id, comments)

# Preprocess data; important for effective sentiment analysis

# Lowercase
comments_df <- comments_df %>%
  mutate(comments = tolower(comments))

# Remove punctuation
comments_df <- comments_df %>%
  mutate(comments = gsub("[[:punct:]]", "", comments))

# Tokenize into one word per row; review_id is kept automatically
text_words <- comments_df %>%
  unnest_tokens(word, comments)

bing <- get_sentiments("bing")

sentiment_bing <- text_words %>%
  inner_join(bing, by = "word")

# if a comment exceeds treshold he will be categorized as either good, neutral or bad
# it depends on the number of positive and negative words/sentiments in the comment
sentiment_per_comment_bing <- sentiment_bing %>%
  group_by(review_id) %>%
  summarise(
    n_positive = sum(sentiment == "positive"),
    n_negative = sum(sentiment == "negative"),
    .groups = "drop"
  ) %>%
  mutate(
    bing_score = case_when(
      n_positive > n_negative ~  1,   # good
      n_positive < n_negative ~ -1,   # bad
      TRUE                    ~  0    # neutral
    )
  )

d.merged <- d.merged %>%
  left_join(sentiment_per_comment_bing, by = "review_id") %>%
  mutate(
    bing_score = if_else(is.na(bing_score), 0L, bing_score)
  )

```

With the merged data in the dataset 'r d.merged' as solid basis is provided for the analysis. The next chapter dives into the data's characteristics, patterns and trends.

# Data Visualization / Story

To answer the research questions, the data must be analysed. In this chapter, different visualizations are examined and their implications discussed. The visualizations are based on the merged dataset from the previous chapter. For most visualizations interim datasets are created, serving as short-term data variables. This allows to keep the *base* dataset clean while loading only the necessary data for visualization and performing data transformations on it if needed.

## Number of listings per room type

First we visualize the number of listings according to the room types. For that it a horizontal barchart displaying the ranking was created. From this visual it is evident that the most listings are rented as entire homes/apartments, followed by private rooms. Hotel and shared rooms make a negligible part of the distribution. Private rooms where guests live in the home of the host are making about one quarter of all listings and the majority of listings take up about three quarters. This means that most listings use entire apartment which take up a lot of living space in comparison to private rooms which are used by the people living in them regardless if the private room is rented or not.


```{r}
library(ggplot2)

# Listing-level counts per room type
room_counts <- d.merged %>% 
  distinct(id, room_type) %>%
  count(room_type, name = "value") %>%
  mutate(
    room_type = factor(
      room_type,
      levels = c(
        "Entire home/apt",
        "Private room",
        "Shared room",
        "Hotel room"
      )
    )
  )

# Horizontal bar chart
ggplot(
  room_counts,
  aes(
    x    = value,
    y    = room_type,
    fill = room_type
  )
) +
  geom_col(width = 0.7) +
  geom_text(
    aes(label = value),
    hjust = -0.1,
    size = 4
  ) +
  scale_fill_manual(
    values = c(
      "Entire home/apt" = "#08519c",
      "Private room"    = "#4292c6",
      "Shared room"     = "#9ecae1",
      "Hotel room"      = "#deebf7"
    ),
    guide = "none"
  ) +
  labs(
    title = "Room types (Listings)",
    x = "Number of listings",
    y = "Room Type"
  ) +
  expand_limits(x = max(room_counts$value) * 1.1) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
  
```

# Neighbourhood group characteristics

In this section the different neighbourhood group characteristics are analyzed. For that the groups are measured according to their total number of hosts and listings aswell as their mean price and mean number of minimum nights of a stay. 

As an overall insights: 
* The numbers of listings and hosts tend to move together meaning groups with many listings also usually have many hosts. This may indicate active Airbnb markets instead of a few large-scale hosts.
* Mean price varies more stronlgy than mean minimum nights. Generally, the minimum nights stay moderate across the city.
* Each group has a different profile suggesting no uniform market.

While we will not go into details of all neighbourhood groups, we will highlight some of them.

Kreis 1, the Old Town, displays higher prices, a moderate number of listings and hosts and shorter minimum stays. This is propably because of its smaller size and likely driven by tourists and business travelers looking for a premium markert. Kreis 2 shows a similar profile.

Some of the groups more far away from the city center like Kreis 12 display a very low interest from hosts to publish listings. Kreis 5 displays as similar pattern but with relatively high prices. Kreis 12 displays its more peripheral and residential character as its being the least active Airbnb area. Kreis 5 with its direction to Schlieren where a more industrial area lies, displays less space for Airbnb's. 

Kreis 3 and 4 with the popular neighbourhood's have profiles with the most hosts and listings indicating the most active market. Which might be also the most affected by possible changes to renting regulations.


```{r}

library(scales)
library(fmsb)

# Aggregate at neighbourhood_group level
radar_raw <- d.merged %>%
  group_by(neighbourhood_group) %>%
  summarise(
    n_hosts        = n_distinct(host_id),
    n_listings     = n_distinct(id),
    mean_price      = mean(price, na.rm = TRUE),
    mean_min_nights = mean(minimum_nights, na.rm = TRUE),
    .groups = "drop"
  )

radar_scaled <- radar_raw %>%
  mutate(
    n_hosts        = rescale(n_hosts,        to = c(0, 100)),
    n_listings     = rescale(n_listings,     to = c(0, 100)),
    mean_price      = rescale(mean_price,      to = c(0, 100)),
    mean_min_nights = rescale(mean_min_nights, to = c(0, 100))
  )

# Set up a 3 x 4 layout for 12 neighbourhood groups
par(mfrow = c(3, 4), mar = c(1, 1, 3, 1))

# Precompute global max and min for the axes (after scaling)
axis_max <- apply(radar_scaled[, -1], 2, max)
axis_min <- apply(radar_scaled[, -1], 2, min)

for (g in radar_scaled$neighbourhood_group) {
  # pick one neighbourhood
  row_g <- radar_scaled %>%
    filter(neighbourhood_group == g) %>%
    select(n_hosts, n_listings, mean_price, mean_min_nights)

  # build the data frame: max row, min row, group row
  df_radar <- rbind(axis_max, axis_min, row_g)
  rownames(df_radar) <- c("max", "min", as.character(g))

  radarchart(
    df_radar,
    axistype = 1,
    pcol     = "#08519c",
    pfcol    = scales::alpha("#4292c6", 0.5),
    plwd     = 2,
    cglcol   = "grey80",
    cglty    = 1,
    axislabcol = "grey40",
    caxislabels = c("0", "25", "50", "75", "100"),
    vlcex    = 0.8
  )
  title(main = g, cex.main = 1)
}

```

 ## Mean price and number of stays per neighbourhood group maps

The prices tend to be higher the closer the neighbourhood group is to the city center. This does not come as a surprise as most people want to minimize their time spent traveling to the city center for their stays. Therefore, listings close to the city center are desired and prices reflect this demand.

Airbnb activit in Zurich is 

```{r}

library(sf)
library(patchwork)

geojson <- st_read("Data/neighbourhoods.geojson")

map_stats <- d.merged %>%
  distinct(id, neighbourhood_group, price, number_of_reviews) %>%
  group_by(neighbourhood_group) %>%
  summarise(
    mean_price  = mean(price, na.rm = TRUE),
    total_stays = sum(number_of_reviews, na.rm = TRUE),
    n_listings  = n(),
    .groups = "drop"
  )

geo_stats <- geojson %>%
  left_join(map_stats, by = "neighbourhood_group")

p_price <- ggplot(geo_stats) +
  geom_sf(aes(fill = mean_price), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(option = "cividis") +
  labs(
    title = "Mean price by neighbourhood group",
    fill  = "Mean price"
  ) +
  theme_minimal()

p_stays <- ggplot(geo_stats) +
  geom_sf(aes(fill = total_stays), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(option = "cividis", trans = "sqrt") +
  labs(
    title = "Number of stays (reviews) by neighbourhood group",
    fill  = "Total stays"
  ) +
  theme_minimal()

p_price + p_stays + plot_annotation(title = "Neighbourhood Groups in Zürich")

```

```{r}
listings <- d.merged %>%
  distinct(id, .keep_all = TRUE)

# Neighbourhood-level stats based on listings
map_stats <- listings %>%
  group_by(neighbourhood) %>%
  summarise(
    mean_price   = mean(price, na.rm = TRUE),
    total_stays  = sum(number_of_reviews, na.rm = TRUE),
    n_listings   = n(),
    .groups = "drop"
  )

# Join stats to neighbourhood geometries
geo_stats <- geojson %>%
  left_join(map_stats, by = "neighbourhood")

# Mean price map (neighbourhood)
p_price <- ggplot(geo_stats) +
  geom_sf(aes(fill = mean_price), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(option = "cividis") +
  labs(
    title = "Mean price by neighbourhood",
    fill  = "Mean price"
  ) +
  theme_minimal()

# Number of stays map (neighbourhood)
p_stays <- ggplot(geo_stats) +
  geom_sf(aes(fill = total_stays), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(option = "cividis", trans = "sqrt") +
  labs(
    title = "Number of stays (reviews) by neighbourhood",
    fill  = "Total stays"
  ) +
  theme_minimal()

# Combine plots
p_price + p_stays + plot_annotation(title = "Neighbourhoods in Zürich")
```

## Interactive map with all listings

```{r}

library(dplyr)
library(leaflet)

airbnb_listings <- d.merged %>%
  distinct(id, .keep_all = TRUE)

# Check that lat/lon look ok (optional)
summary(airbnb_listings$latitude)
summary(airbnb_listings$longitude)

# Basic leaflet map with basemap + all points
leaflet(airbnb_listings) %>%
  # tile layer with streets etc. 
  addProviderTiles(providers$CartoDB.Positron) %>%
  
  # each Airbnb as a point with information
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 3,
    stroke = FALSE,
    fillOpacity = 0.6,
    color = "#FF5733",
    clusterOptions = markerClusterOptions(),
    popup = ~paste0(
      "<b>", name, "</b><br/>",
      "Room type: ", room_type, "<br/>",
      "Price: ", price, " CHF<br/>",
      "Neighbourhood: ", neighbourhood
    )
  )


```

# Reviews analysis on maps

```{r}

neg_stats <- d.merged %>%
  filter(bing_score == -1) %>%
  group_by(neighbourhood) %>%
  summarise(
    n_negative_reviews = n(),
    .groups = "drop"
  )

all_stats <- d.merged %>%
  group_by(neighbourhood) %>%
  summarise(
    n_reviews = n(),
    .groups = "drop"
  )

neg_stats <- neg_stats %>%
  left_join(all_stats, by = "neighbourhood") %>%
  mutate(
    share_negative = n_negative_reviews / n_reviews
  )

geo_neg <- geojson %>%
  left_join(neg_stats, by = "neighbourhood")

p_count <- ggplot(geo_neg) +
  geom_sf(aes(fill = n_negative_reviews), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(
    option = "cividis",
    na.value = "grey90",
    name = "Negative reviews"
  ) +
  labs(
    title = "Number of negative reviews per neighbourhood",
    subtitle = "Lighter areas have more negative reviews"
  ) +
  theme_minimal()

p_share <- ggplot(geo_neg) +
  geom_sf(aes(fill = share_negative), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(
    option = "cividis",
    na.value = "grey90",
    labels = scales::percent_format(accuracy = 1),
    name = "% negative"
  ) +
  labs(
    title = "Share of negative reviews per neighbourhood",
    subtitle = "Fraction of reviews with bing_score = -1"
  ) +
  theme_minimal()

p_count + p_share + plot_annotation(
  title = "Negative Airbnb reviews in Zürich by neighbourhood"
)


```

## Distribution of minimumg nights

```{r}

interim <- d.merged %>%
  distinct(id, .keep_all = TRUE)

ggplot(
  interim %>% filter(minimum_nights <= 35),
  aes(x = minimum_nights)
) +
  geom_histogram(
    binwidth = 1,
    fill = "#3182bd",
    alpha = 0.8,
    boundary = 0
  ) +
  scale_x_continuous(breaks = seq(0, 35, 2)) +
  labs(
    title = "Distribution of minimum nights (1–35, listing level)",
    x = "Minimum nights",
    y = "Number of listings"
  ) +
  theme_minimal()

```

## Minimum nights per neighbourhood map

```{r}

library(leaflet)
library(scales) 

listings <- d.merged %>%
  distinct(id, .keep_all = TRUE)

# mean minimum_nights per neighbourhood
neigh_stats <- listings %>%
  group_by(neighbourhood) %>%
  summarise(
    mean_minimum_nights = mean(minimum_nights, na.rm = TRUE),
    n_listings         = n(),
    .groups = "drop"
  )

# Join to neighbourhood polygons and transform to WGS84 (lat/lon)
geo_min <- geojson %>%
  left_join(neigh_stats, by = "neighbourhood")

geo_min_4326 <- st_transform(geo_min, 4326)

# Point per neighbourhood (for bubbles)
neigh_pts <- st_point_on_surface(geo_min_4326)
coords    <- st_coordinates(neigh_pts)

pts_df <- cbind(
  st_drop_geometry(neigh_pts),
  lon = coords[, "X"],
  lat = coords[, "Y"]
)

# Colour + size
pal <- colorNumeric("cividis", domain = pts_df$mean_minimum_nights)
radius_scaled <- rescale(pts_df$mean_minimum_nights, to = c(4, 16))

# Interactive leaflet map with neighbourhood borders
leaflet(geo_min_4326) %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik) %>%
  
  # neighbourhood outlines 
  addPolygons(
    color         = "#444444",     # border colour
    weight        = 1.2,           # thickness of line
    opacity       = 0.9,
    fill          = FALSE,         # just outlines
    highlightOptions = highlightOptions(
      color = "red",
      weight = 2,
      bringToFront = TRUE
    ),
    label = ~neighbourhood         # hover label on polygons 
  ) %>%
  
  # bubbles per neighbourhood
  addCircleMarkers(
    data        = pts_df,
    lng         = ~lon,
    lat         = ~lat,
    radius      = radius_scaled,
    color       = "#ffffff",
    weight      = 1,
    fillColor   = ~pal(mean_minimum_nights),
    fillOpacity = 0.85,
    popup = ~paste0(
      "<b>", neighbourhood, "</b><br>",
      "Mean minimum nights: ", round(mean_minimum_nights, 1), "<br>",
      "Listings: ", n_listings
    )
  ) %>%
  
  addLegend(
    pal      = pal,
    values   = ~mean_minimum_nights,
    position = "bottomright",
    title    = "Mean minimum nights"
  )

```





```{r}

min_nights <- d.merged %>%
  distinct(id, minimum_nights) %>%
  count(minimum_nights) %>%
  arrange(minimum_nights) %>%
  mutate(
    share = n / sum(n),
    cum_share = cumsum(share)
  )

print(n = 53, min_nights)

```




# Model

# Conclusions

# Potential Limitations

# GenAI

Ideas of questions to answer:

**What have you used it for?**

**For what kind of tasks turned they out to be helpful?**

**How do you check correctness of the answer?**

**What did not work?**

# Requirements

We cross them as soon as fulfilled:

Two or more data sets which need to be joined/merged

**First data set:**

At least few hundred of observations

At least dozen variables

Numeric and categorical variables

Dates ( YYYY-MM-DD) or geographic locations, both are even better

**Second data set:**

Can be simple

Join the two data sets and highlight in your report where you have done
this, showing the line of code where the joining is performed

Ideally different formats and different sources

**Content** Comprehensive analysis which we will present to a hypothetical
client

In depth examination of problem being adressed

PDF or HTML with Rmarkdwon or quarto

Clear and comprehensible to all readers

**Prepare the data for the analysis** Cleaned before they are merged and
analysis begins

You can hide code but highlight in the text where the data is merged

Please explain in words (not as code) what you have done for data
preparation

**Visualize the data appropiately**
Analyse data using summary statistics

Graphs

*Recommendations for plots*
No pie charts

Alpha level to add transparency to points /lines

Not to many stacked bar plots, or not at all (instead maybe line plots)

Show entire range of values, without zooming too much

Same scale on axes

Shall all observations and infos, not just mean and standard deviation

Not too many dimensions in a graph

**Fit the model**
Fit a model and show result + interpet it

Produce graphs (e.g. predictions or residual diagnostics) for your model fits

Possibility: compare several models via CV

**Chapter of choice**
Use new package that was not mentioned in the main part of the course (cant be data
manipulation, visualisation using ggplot2, regular expressions and reporting using Rmarkdown does not count and perform a task that we have not discussed in the course

No new statistical or machine learning method

Method to prepare or display data

**Generative AI**
Quarter to half a page about which tools you have used

What have you used it for? For what kind of tasks turned they out to be
helpful?

How do you check correctness of the answer?

What did not work?

**Sell the story**
Complete, easy to read, thread (what are you doing, why, conclusions, interpretations, overview . . . )

• Source of the data.
• Objectives and hypotheses of the analysis.
• Interpretation of each step.
• Conclusions drawn from the analysis.
• Potential limitations of the study.

**Comments**
Comment a lot

Check coding style

**Additional requirements**
a readme.txt file;

structured folders (e.g. “Data”, “Scripts”, . . . )

the data set analysed

the Rmarkdown file (.Rmd) (or quarto file)

the PDF or HTML output file

not more than 25 pages

